---
title: "Craigslist apartments"
author: "Sue Chew"
date: Nov 21, 2005
output: pdf_document
---

# Overview
We wish to build an application that helps a user find out the price and size of the apartment that the user is likely to get, given a city and the desired number of bedrooms.

To do this we are going to obtain recent data from craigslist's "apt/housing" for a few United States cities and use it to train a model to predict price and size.

# Data processing
It may not be exactly enough data, but lets try doing this with just the 100 most recent posts from each city, as obtained on the evening of November 20 2015.
`get_craigslist_data.R` was used to download the data, and it stored into `data/citiesdf.rds`

Print summary.
```{r}
citiesdf <- readRDS('data/citiesdf.rds')
str(citiesdf)
summary(citiesdf)
```

# Exploratory plots
```{r}
library(ggplot2)
ggplot(citiesdf, aes(y=price, x=city)) + geom_violin() +
    labs(title="Price vs City")

ggplot(citiesdf, aes(x=sqft, y=price, color=city)) + geom_point() +
    labs(title="Price vs Sqft by City")

ggplot(citiesdf, aes(x=bedrooms, fill=as.factor(bedrooms))) + geom_histogram() +
    coord_flip() + facet_wrap(~ city) +
    labs(title="Number of Bedrooms per City")
```


# Model training
`train_models.R` was used to train two models, which are as follows.

| model formula | saved to file |
|---------------|---------------|
| price ~ city + bedrooms | data/pricemod.rds |
| sqft ~ city + bedrooms | data/sqftmod.rds |

```{r}
pricemod <- readRDS('data/pricemod.rds')
summary(pricemod)

sqftmod <- readRDS('data/sqftmod.rds')
summary(sqftmod)
```

# Check out some predictions
```{r}
test <- data.frame(
    city = rep(c('chicago', 'honolulu', 'kansascity', 'losangeles', 'newyork', 'philadelphia', 'sfbay', 'seattle'), 3),
    bedrooms = rep(c(1,2,3)))
test <- test[order(test$city, test$bedrooms),]

# Predict price (USD) and size (sqft).
cbind(test,
      predictedPrice=predict(pricemod, newdata = test),
      predictedSize=predict(sqftmod, newdata = test))
```

