---
title: "Craigslist apartments"
author: "Sue Chew"
date: "Nov 21, 2005"
output: pdf_document
---

# Overview
We wish to build an application that helps a user find out the price and size of the apartment that the user is likely to get, given a city and the desired number of bedrooms.

To do this we are going to obtain recent data from craigslist's "apt/housing" for a few United States cities and use it to train a model to predict price and size.

# Data processing
It may not be exactly enough data, but lets try doing this with just the 100 most recent posts from each city.
`get_craigslist_data.R` was used to download the data, and it stored into `data/citiesdf.rds`

Print summary.
```{r}
citiesdf <- readRDS('data/citiesdf.rds')
str(citiesdf)
summary(citiesdf)
```

# Exploratory plots

```{r, warning=FALSE, message=FALSE}
library(Hmisc)
library(dplyr)
library(ggplot2)
```

```{r}
ggplot(citiesdf, aes(y=price, x=city)) + geom_violin() +
    labs(title="Price vs City")

ggplot(citiesdf, aes(x=sqft, y=price, color=city)) + geom_point() +
    labs(title="Price vs Sqft by City")

ggplot(citiesdf, aes(x=bedrooms, fill=as.factor(bedrooms))) + geom_histogram() +
    coord_flip() + facet_wrap(~ city) +
    labs(title="Number of Bedrooms per City")



```

Since sqft is not always available, imputing missing values so that we can calculate price per square foot.
```{r}
citiesdf.ar <- aregImpute(~ bedrooms + sqft, data=citiesdf, n.impute=1)
# Create a new data frame with the imputed sqft.
citiesdf.imp <- citiesdf
citiesdf.imp$sqft[as.numeric(row.names(citiesdf.ar$imputed$sqft))] <-
    citiesdf.ar$imputed$sqft
summcity <- summarise(group_by(citiesdf.imp, city),
                      avePricePerSqft = sum(price) / sum(sqft))
ggplot(summcity, aes(x=city, y=avePricePerSqft)) + geom_point(size=3) +
    ggtitle('Average price per sqft') +
    geom_hline(aes(yintercept=mean(summcity$avePricePerSqft)), linetype="dashed")
```

# Model training
`train_models.R` was used to train two models, which are as follows.

| model formula | saved to file |
|---------------|---------------|
| price ~ city + bedrooms | data/pricemod.rds |
| sqft ~ city + bedrooms | data/sqftmod.rds |

```{r}
pricemod <- readRDS('data/pricemod.rds')
summary(pricemod)

sqftmod <- readRDS('data/sqftmod.rds')
summary(sqftmod)
```

# Check out some predictions
```{r}
test <- data.frame(
    city = rep(c('chicago', 'honolulu', 'kansascity', 'losangeles', 'newyork',
                 'philadelphia', 'sfbay', 'seattle'), 4),
    bedrooms = rep(c(0,1,2,3), each=8))
test <- test[order(test$city, test$bedrooms),]

# Predict price (USD) and size (sqft).
predictedPrice <- predict(pricemod, newdata = test)
predictedSize <- predict(sqftmod, newdata = test)

pred <- cbind(test,
      predictedPrice = round(predictedPrice),
      predictedSize = round(predictedSize),
      predictedPricePerSqft = round(predictedPrice/predictedSize, 1))

print(pred, row.names=FALSE)

ggplot(pred, aes(x=bedrooms, y=predictedPrice)) + geom_point() +
    geom_line() + facet_wrap( ~ city, scales="free_x") +
    ggtitle("Predicted price vs bedrooms by city")

ggplot(pred, aes(x=bedrooms, y=predictedSize)) + geom_point() +
    geom_line() + facet_wrap( ~ city, scales="free_x") +
    ggtitle("Predicted size (sqft) vs bedrooms by city")
```

